The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-03-08 src/ms.tex
Apply suggestions from code review

% Summarize funding sources (Moore, NASA) and amounts and what this has been
% used for.

With the goal of establishing a sustainable financial model for the Astropy
Project, the Coordination Committee under the new governance charter (see \S
\ref{sec:project-governance}) established a standing Finance committee.
This committee oversee the planning and allocation of finances on behalf
of the project. The Astropy Project accepts funds from institutions as well as
individuals through NUMFOCUS\footnote{NUMFOCUS is a 501(c)(3) nonprofit that
supports and promotes world-class, innovative, open source scientific computing.}.
Previously, no direct financial support was available for the project development,
and NUMFOCUS covered most of the incurred operational costs. However, transitioning
to long-term financial stability meant having an influx of funds in the form of
institutional support, and hence the Astropy Project applied for and successfully
acquired four grants. These grants are administered by NumFOCUS on behalf of the
project, providing a ``neutral'' space not tied to any particular astronomy
institution. The two major ones are briefly summarized below.

\begin{itemize}
    \item Moore: The Gordon and Betty Moore Foundation awarded the Astropy
    project a $\sim$\$900k (US) grant in 2019 for the proposal ``Sustaining and
    Growing the Astropy Project". Over a three-year term, this grant supported
    sustenance of the project by providing funds for (1) the development and
    maintenance of the \astropypkg and its infrastructure to existing project
    members and targeted hires, (2) the transition of long-term users to
    contributors and maintainers through mentorship efforts, (3) the formalization
    of a governance structure (see \S \ref{sec:project-governance}), and (4) the
    improvement of equity, diversity, and inclusion efforts in the project (see \S
    \ref{sec:project-ide}). Some of the money supported travel for meetings,
    conferences, and workshops. This generous grant helped reduce reliance on
    volunteer-driven maintenance of the openly-developed package and paved
    the way to acquire funding from federal agencies: a big milestone for the
    success of the project.

    \item NASA: A successful proposal to the NASA ROSES-2020
    call\footnote{http://solicitation.nasaprs.com/ROSES2020}, section E.7
    (Support for Open Source Tools, Frameworks, and Libraries), granted the
    Astropy Project $\sim$\$600k (US) to support its work. Funded work for a
    term of three years includes (1) project-wide infrastructure maintenance and
    improvements, (2) targeted work on areas of the \astropy core and
    coordinated packages that need support, (3) enhancements to the Learn
    Astropy ecosystem and educational materials (see \S \ref{sec:learn}),
    and (4) support for the \astropy affiliated packages. These goals
    support the sustenance of the project by encouraging further
    development of the package ecosystem and community engagement.
\end{itemize}

In addition to the two grants described above, the project benefited from
a Gemini Observatory contract under its Science User Support Department,
which awarded funds for development work that supports both the Astropy and
DRAGONS projects. The project also received financial support from the Dunlap
Seed Funding program at the University of Toronto to develop educational
resources for scientific software packaging within the Learn Astropy
framework. All these successful proposals show that the Astropy project
is vital for the astronomical infrastructure and community at large, and
strengthen its promising and sustainable future.

% ADD text for the new finance APE once it has a doi

@@

Overview of roadmap and some highlights.

>>>>

Summarize funding sources (Moore, NASA) and amounts and what this has been used
for.

@@

The Astropy Roadmap\footnote{\url{https://github.com/astropy/astropy-project/blob/main/roadmap/roadmap.md}} is a document in the astropy-project repository that captures high level actionable items that the Astropy project aims to undertake to improve the health and stability of the  project. It is a static document that is revisited regularly at the Astropy coordination meetings, to keep track of progress and write new versions as needed. There is a related project board linked to the Roadmap document, that holds specific issues and efforts related to Roadmap items. The project board is a living document that is continually updated as work is planned, assigned, and completed.

All items in the Roadmap have been agreed to be priorities for the Astropy Project, and are color-coded based on resources (both time/effort and developers/expertise) needed to complete the item. Green items are well underway, have sufficient resources/support and a plan in place for completion. Orange items are well defined, and work for aquiring sufficient resources underway. Red items do not yet have a plan for implementation and need more resources. 
\todo{addition}{Add some examples of high priority items from the roadmap in a new paragraph here}
The Astropy Roadmap originates from the March 2021 Astropy Coordination meeting where it was first drafted before being handed off to a newly formed Astropy Roadmap working group for completion. The Roadmap in its currant form was adopted via pull request in December 2021.
